---
title: "Doing Descriptive Statistics in R"
author: "Anastasia Molotova"
date: "12/05/2020"
output: 
  html_document:
    theme: spacelab
    highlight: monochrome
    toc: true
    toc_float: true
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, error = FALSE)
```

# About this project

The aim of this work is to give a brief introduction into R, descriptive statistics and the best packages for these purposes. Claim it not as an essay but as a tutorial for R-beginners and experience sharing.  

# Data

First of all, I chose the data to perform the analysis on. For this project I will use the build-in dataset `Salaries` from package `carData`, because there are only 6 observations with 3 factors and 3 integers (I believe it is enough for now). The data is "nine-month academic salary in a college in the U.S" (from `help`).

```{r}
library(carData)
#?Salaries
data = Salaries
str(data)
```

# Evaluation criteria

In order to choose the best tool for myself and evaluate different approaches somehow, I developed criteria:

1. **SCI**: completeness of information in summary output (aka *research requirements*), if it is enough or not information to make some advanced conclusions *[1]
2. **RCI**: completeness of information in regression output (aka *research requirements*), if it is enough or not information to make some advanced conclusions *[1]
3. **SC**: clearness of summary output (aka *publishing standards*), if is clear enough, not overloaded *[1]
4. **RC**: clearness of regression output (aka *publishing standards*), if is clear enough, not overloaded *[1]

 *[1] the aim of this project is to identify a toolset than will be on the "golden middle" between "a lot" and "nothing at all" for comfortable and universal usage

My additional criteria: 

3. **SIMPLICITY**: either it is easy to perform, easy to use, easy to code
4. **GRMD**: either looks good in rmd
5. **GHTML**: either looks good in html
6. **GROUPING**: available option of `group by` grouping variable in summary statistics
7. **TITLE**: either we can add title to the table (0.5 points for summary as well as for models, 1 overall)

As for competitors, I will compare tools provided by `base R`, `skimr`, `stargazer` and `summarytools` step-by-step, giving some comments and writing ideas behind grading. Those packages were used because I have already tested them on some projects and they were the best!

*Author's note: all grading is my personal feeling and experience and might not match someone's opinion. Another point is that you might need both rmd and html files to observe the difference in the output according to the criteria. All mistakes are my own, but if I did not find some options and said "there is no title", it means either my blindness is getting worse, or the author might think of making title part more noticeable.*

In the final part I will create a leaderboard with all these criteria as a binary outcome (either 1 if fine, 0 if not) and write a conclusion.

# 1 -- base R: starterpack

## Examples

Basic `summary` which is known by any statistician-beginner:

```{r}
s = summary(data)
s
```

Usual output with 6 main statistics for "numerics" and counted factors, but we can try to make it better with `kable` package:

```{r}
library(tidyverse)
library(kableExtra)
a = s %>%
  kbl() %>%
  kable_styling()
a
```

Well, not my best try so far, looks bad (especially NAs).

Also, there is a function `by` which produce statistics by group:

```{r}
by(data, data$sex, summary)
```

As for linear regression, now I will perform a model and then just test the output. The model will predict salary, according to the years since PHD was completed, years of service and gender. 

```{r}
model <- lm(salary ~ yrs.since.phd + yrs.service + sex, data = data)
```

`Summary` for the model:

```{r}
summary(model)
```

Looks nice to me.

## Grading

|        | SCI | RCI |  SC |  RC | SIMPLICITY | GRMD | GHTML | GROUPING |  TITLE   |
|-------:|:---:|:---:|:---:|:---:|:----------:|:----:|:-----:|:--------:|:--------:|
| base R |  0  |  1  |  1  |  1  |     1      |  1   |    0  |     1    |   0      |

Comments on grading:

> SCI: output describe only 6 main statistics, I prefer to see something else. It is easy to perform quantiles or other stats, but requires additional code (0/1)

> RCI: regression looks good to me (1/1)

> SC: clear enough (1/1)

> RC: all neccesary info such adjusted R-squared and significance codes are here (1/1)

> SIMPLICITY: easy enough (1/1)

> GRMD: usual rmarkdown output (1/1)

> GHTML: no difference between html and rmd (0/1)

> GROUPING: there is an option for grouping stats (1/1)

> TITLE: no option of changing the title (0/1)

Overall score: **6/9**
 

# 2 -- skimr

Official [Github](https://github.com/ropensci/skimr)

## Examples

Basic summary:

```{r}
library(skimr)

skim(data)
```

Summary by group:

```{r}
group_by(data, sex) %>%
 skim()
```

## Grading

|        | SCI | RCI |  SC |  RC | SIMPLICITY | GRMD | GHTML | GROUPING |  TITLE   |
|-------:|:---:|:---:|:---:|:---:|:----------:|:----:|:-----:|:--------:|:--------:|
| skimr  |  1  |  0  |  1  |  0  |     1      |  1   |    0  |     1    |     0    |

Comments on grading:

> SCI: output describe main statistics with NA and distribution! Love it (1/1)

> RCI: this package does not provide any model output (0/1)

> SC: clear enough (1/1)

> RC: no regression output (0/1)

> SIMPLICITY: easy enough (1/1)

> GRMD: usual rmarkdown output (1/1)

> GHTML: unfortunately, there are some troubles with histogram that i love in html (strange numbers instead of beauty) (0/1)

> GROUPING: there is an option for grouping stats (1/1)

> TITLE: no information on title in official documentation (0/1)

Overall score: **5/9**

# 3 -- stargazer

Official [CRAN page](https://cran.r-project.org/web/packages/stargazer/stargazer.pdf)

## Examples

Basic summary:

```{r}
library(stargazer)
stargazer(data)
```

Wow! You need to write additional argument in order to avoid such a mess:

```{r}
stargazer(data, type = "text", title = "Summary Statistics")
```

Now looks better. It is also possible to create LaTeX and html outputs, but we have to change rmd or chunk accordingly:

```{r, results ='asis', message = FALSE}
stargazer(data, type = "html", title = "Summary Statistics")
```

As for model output, there is two options as well. First one is a "text" option:

```{r}
stargazer(model, type = "text", title = "Regression results")
```


Also, special setting is required for beautiful linear regression output:

```{r, results ='asis', message = FALSE}
stargazer(model, type = "html", title = "Regression results")
```

Moreover, this package provides different journals styles. Previously, there was a default option (see `stargazer: list of supported styles` in help), but here is a style of *American Journal of Sociology* as an example:

```{r, results ='asis', message = FALSE}
stargazer(model, type = "html", title = "Regression results", style = "ajs")
```

You can see the difference in style, especially in p-value appearance.

## Grading

|         | SCI | RCI |  SC |  RC | SIMPLICITY | GRMD | GHTML | GROUPING |  TITLE   |
|--------:|:---:|:---:|:---:|:---:|:----------:|:----:|:-----:|:--------:|:--------:|
|stargazer|  0  |  1  |  1  |  2  |     0      |  1   |    0  |     0    |   1      |

Comments on grading:

> SCI: output describe few main statistics, there are no obvious ways to expand this list (0/1)

> RCI: all main information is provided (1/1)

> SC: fine, you can just take a screenshot and go publish the results (1/1)

> RC: I will put 2 points here! Models are the best part of this package: you basically do not need to take care of the output's beauty, it will produce even special style! I used it several times in my projects and term paper, and fall in love with this package since then!  (2/1)

> SIMPLICITY: you have to know how to deal with output paramentrs or google it (0/1)

> GRMD: basically usual rmarkdown output (1/1)

> GHTML: awesome knitted html! (1/1)

> GROUPING: I did not find a word about grouping option (0/1)

> TITLE: there is an option of chanding/adding title into the output in both summary or model (1/1)

Overall score: **6/9**

# 4 -- summarytools

Official [Github](https://github.com/dcomtois/summarytools)

## Examples

Overall statistics:

```{r}
library(summarytools)
print(dfSummary(data, graph.magnif = 0.75), method = 'render')
```

Summary by group:

```{r}
data %>% group_by(sex) %>% dfSummary(plain.ascii = FALSE, style = "grid", 
          graph.magnif = 0.75, valid.col = FALSE, tmp.img.dir = "/tmp")
```

Some ugly monster, but, according to the official Github, everything must work. Must be my mistake, but I didn't find how to deal with it.

Summary for numerical only:

```{r}
descr(data) 
```

We can describe one exact numeric variable by another categorical: 

```{r}
with(data, stby(salary, sex, descr))
```


## Grading

|            | SCI | RCI |  SC |  RC | SIMPLICITY | GRMD | GHTML | GROUPING | TITLE |
|-----------:|:---:|:---:|:---:|:---:|:----------:|:----:|:-----:|:--------:|:-----:|
|summarytools|  1  |  0  |  1  |  0  |     0      |  0.5 |    1  |     1    |   0   |

Comments on grading:

> SCI: powerful statistics, first time when i saw frequency and such a detailed report on variables type (1/1)

> RCI: no approaches for model output is provided (0/1)

> SC: nice knitted tables with a little effort  (1/1)

> RC: again, no options for model output (0/1)

> SIMPLICITY: using this summary requires "effort" and some knowledge (0/1)

> GRMD: all the functions working as usual output (except main dfSummary) (0.5/1)

> GHTML: dfSummary with a little effort knits to a beautiful html table! (1/1)

> GROUPING: several options for grouping (1/1)

> TITLE: there is no need to change it, but either no option for this (0/1)

Overall score: **4,5/9**

# Leaderboard and conclusion

|            | SCI | RCI |  SC |  RC | SIMPLICITY | GRMD | GHTML | GROUPING |  TITLE   |OVERALL|
|-----------:|:---:|:---:|:---:|:---:|:----------:|:----:|:-----:|:--------:|:--------:|:-----:|
| base R     |  0  |  1  |  1  |  1  |     1      |  1   |    0  |     1    |   0      | 6     |
| skimr      |  1  |  0  |  1  |  0  |     1      |  1   |    0  |     1    |   0      | 5     |
|stargazer   |  0  |  1  |  1  |  2  |     0      |  1   |    0  |     0    |   1      | 6     |
|summarytools|  1  |  0  |  1  |  0  |     0      |  0.5 |    1  |     1    |   0      | 4,5   |

Overall, I would recommend `base R` for the beginners, since it is user-friendly, and basically has the main statistics information.
`skimr` and `summarytools` provide powerful descriptive statistics, while `stargazer` is more suitable for models' output.

Personally, I use `base R` for fast stats checking in console for last three years. For my term paper I needed more clear output, preferably pretty knitted in html and used `stargazer` for this. 

# Afterword

Another nice package that is loved by classmates and me is [`DataExplorer`](https://www.rdocumentation.org/packages/DataExplorer/versions/0.8.2), but for unknown reasons there might be some troubles with its installation. Moreover, according to the author's [Github page](https://github.com/boxuancui/DataExplorer) it is not published in CRAN. Still this package provide powerful EDA with pretty plots.

