---
title: "Cluster Times HE"
author: "Anastasia Molotova"
date: "11/19/2020"
output: 
  html_document:
    theme: paper
    highlight: monochrome
    toc: true
    toc_float: false
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```

```{r}
library(readr)
df <- read_csv("~/df.csv")
```

# Research idea

First of all, I would like to introduce my own research question on which I will be answering in this project: 

> "I want to find a University where I will receive a large salary after graduation". Another limitation is my desire to study either political science or sociology.

I will focus on choosing variables first and EDA then.

# Data preparation

So, this is mostly market-oriented project, therefore I want only `income` variable and score of teaching (we still want to study in the best uni on the planet!):

```{r}
#colnames(df)

library(dplyr)
df = df %>% 
  select(-scores_research, -scores_research_rank, -scores_citations, -scores_citations_rank, -stats_pc_intl_students, -scores_international_outlook, -scores_overall)
```

Deleting also some half-metadata info and ranks (prefer working with scores), also deleted location (applicants from all over the world and ready to move):

```{r}
df = df %>% 
  select(-rank_order, -rank, -record_type, -url, -member_level, -aliases, -closed, -apply_link, -ref_link, -scores_overall_rank, -scores_teaching_rank, -scores_industry_income_rank, -scores_international_outlook_rank, -stats_female_male_ratio, -location, -stats_female_share, -stats_number_students, -stats_student_staff_ratio)
```

```{r}
#df$scores_overall = strtrim(df$scores_overall, 4) #get rid of "50.6–54.2"
#df$scores_overall = as.numeric(df$scores_overall)
```

I will also create dummies for `subject` in order to get some interesting clusters:

```{r dummies for subjects}
library(dplyr)
library(tidyr)

df_subj = df %>% 
  mutate(subjects_offered = strsplit(as.character(subjects_offered), ",")) %>%
  unnest(subjects_offered)
df_subj$subjects_offered = trimws(df_subj$subjects_offered)

library(fastDummies)
data <- dummy_cols(df_subj, select_columns = 'subjects_offered')
data = data %>% select(-subjects_offered)


data[,c(10:45)] <- lapply(data[,c(4:39)] , factor)
#data[, c(10:45)] <- as.factor(data[, c(10:45)])

```

# EDA

```{r}
library(DataExplorer)
```

```{r}
plot_intro(data)
```

These are our dummy variables in dataset:

```{r}
plot_bar(data)
```

Correlations first:

```{r}
cor_num = data[, c(2:3)]

require(heatmaply)
heatmaply(
  cor(normalize(cor_num)),
  xlab = "Features",
  ylab = "Cars"
)
```

Overall, correlations looks good for now.

```{r}
#Standardizing numeric variables:
#data[,c(2:5, 7:9)] <- lapply(data[,c(2:5, 7:9)] , scale)
```

Here are our numeric variables:

```{r}
plot_histogram(data)
```

# Distance metric 

```{r}
#categories = data %>% select(nid, name)
data <- data[-c(41:45)]
#data = distinct(data)
data = data %>% select(-nid)
data = data[,c(1:3, 35,37)]
data = distinct(data)



library(cluster)
gower_dist <- daisy(data[ , -1],
                    metric = "gower",
                    type = list(logratio = 3)) 

summary(gower_dist)
```

Fine, types by `daisy` are true: I have 2 factor variables in the end and 2 numerical.

# how many clusters must be here

```{r}
sil_width <- c(NA)

for(i in 2:10){
  pam_fit <- pam(gower_dist,
                 diss = TRUE,
                 k = i)
  sil_width[i] <- pam_fit$silinfo$avg.width
}
plot(1:10, sil_width,
     xlab = "Number of clusters", xaxt='n',
     ylab = "Silhouette Width",
     ylim = c(0,1))
axis(1, at = seq(2, 10, by = 1), las=2)
lines(1:10, sil_width)
```

The highest value is 3, so we must choose this number of clusters.

#PAM clustering

```{r}
pam_fit <- pam(gower_dist, diss = TRUE, k = 3)
pam_fit$data <- data[ , -1]
```

```{r}
pam_results <- data %>%
  dplyr::select(-name) %>%
  mutate(cluster = pam_fit$clustering) %>%
  group_by(cluster) %>%
  do(the_summary = summary(.))
pam_results$the_summary
```

*Cluster description on PAM method*: 

- Cluster 1 represents all the universities with no political science/sociology

- Cluster 2 represents sociological universities and teachers there are better

- Cluster 3 represents political science and income is higher

From our customer's point of view the better option is political sciences and higher income (cluster 3), and best choices will be:

```{r}
data[pam_fit$medoids, ]
```

```{r}
set.seed(99)
library(Rtsne)
tsne_obj <- Rtsne(gower_dist, is_distance = TRUE)

tsne_data <- tsne_obj$Y %>% # $Y is a matrix containing the new representations for the objects
  data.frame() %>%
  setNames(c("X", "Y")) %>%
  mutate(cluster = factor(pam_fit$clustering),
         name = data$name)
library(ggplot2)
ggplot(aes(x = X, y = Y), data = tsne_data) +
  geom_point(aes(color = cluster))
```

Exactly three clusters on the plot!